Exploratory analysis
========================================================

### Load libraries

```{r}
```

Download the data, read the data in and save the raw data as an RDA file

```{r loansData, cache=TRUE}
setwd("/dvt/workspace/r-data-analysis-course/assignment-1/work")
getwd()
download.file("https://spark-public.s3.amazonaws.com/dataanalysis/loansData.csv",destfile="./data/raw/loansData.csv",method="curl")
dateDownloaded <- date()
dateDownloaded
loansData <- read.csv('./data/raw/loansData.csv')
save(loansData,dateDownloaded,file="./data/raw/loansData.rda")
```

### Fix incorrect types
```{r dependson="loansData"}
loansData$Interest.Rate <- sapply(loansData$Interest.Rate, function(x) as.numeric(sub("%", "", x)))
loansData$Debt.To.Income.Ratio <- sapply(loansData$Debt.To.Income.Ratio, function(x) as.numeric(sub("%", "", x)))
```

### Look at the data set

```{r}
str(loansData)
sapply(loansData, summary)
```

### Find out about missing values
```{r, dependson="loansData"}
sum(is.na(loansData))
loansData <- loansData[complete.cases(loansData),]
```

### Explore the suspiscious columns
```{r, dependson="loansData"}
hd <- hist(loansData$Monthly.Income, breaks=100)
print(hd$counts)
# Monthly needs transformation
hd <- hist(log(loansData$Monthly.Income), breaks=100)
max(loansData$Monthly.Income, na.rm=T)
```
There is one monthly income which seems extremely high, probably a mistake?

```{r}
suspi <- subset(loansData, Monthly.Income == max(loansData$Monthly.Income, na.rm=T))
print(suspi)
suspi$Monthly.Income / suspi$Amount.Funded.By.Investors
```
The percentage of gross income that goes into paying the debt is small, which indicates that the person may really have a big income. Though some others do have smaller ratios.


## Exploratory analysis

### Make some univariate plots/summaries

```{r, dependson="loansData"}
loansData$FICO <- sapply(as.character(loansData$FICO.Range), function(s) as.numeric(substr(s, 0, 3)))

# interest rate seems interesting
hist(loansData$Interest.Rate, breaks=100)
# as if there were two distributions...
hist(log10(loansData$Interest.Rate), breaks = 100, prob=T)

# this is interesting, there is a 'required' minimum
hist(loansData$FICO, breaks=100)
# log transformation doesnt change it (so it is cutoff)
hist(log(loansData$FICO), breaks=100)

# is this increasing lineary? it seems like...
quantile(loansData$FICO, na.rm=T, probs=seq(0, 1, 0.05))

# investors obviously prefer borrowers who don't have too
# high ratio of indebtness, but it is not too strong
hist(loansData$Debt.To.Income.Ratio)

# and few credit lines
hist(loansData$Open.CREDIT.Lines, breaks=100)
hist(as.numeric(cut(loansData$Open.CREDIT.Lines, breaks=seq(0,30,1))))

# what was requrested was granted, hmm....
par(mfcol=c(2,1))
par(mar=c(2.5,2.5,1.6,1.1),mgp=c(1.5,.5,0))
hist(loansData$Amount.Requested, breaks=100)
hist(loansData$Amount.Funded.By.Investors, breaks=100)
par(mfcol=c(1,1))

# my understanding is that only borrowers who do not
# already own too much, are welcomed
# TODO: look at the relationship between CB ~ FICO
quantile(loansData$Revolving.CREDIT.Balance, na.rm=T, probs=seq(0, 1, 0.05))
hist(loansData$Revolving.CREDIT.Balance, breaks=100)
# needs transformation...
hist(log(loansData$Revolving.CREDIT.Balance), breaks=100)

# TODO: has employment have any effect on the interest?
hist(as.numeric(loansData$Employment.Length), breaks=100)



quantile(loansData$Interest.Rate)
quantile(loansData$Monthly.Income, na.rm=T, probs=seq(0, 1, 0.05))


```

So it seems that FICO must be > 650 for the loan to be approved. And we have got some terms to check for relations


### Check out relationship with Interest Rate (IR)
```{r, dependson="loansData"}
ir <- loansData$Interest.Rate
par(las=3)
par(mar=c(6,2,1,1))
boxplot(ir ~ loansData$Loan.Purpose )
boxplot(ir ~ loansData$Home.Ownership)

# a linear relationn between IR an FICO?
boxplot(ir ~ loansData$FICO.Range)

# no relation between IR and Employment length
boxplot(ir ~ loansData$Employment.Length)

# IR clearly depends on the length (which is logical, but
# it is interesting that shorter terms have lower IR)
boxplot(ir ~ loansData$Loan.Length)

# am i seeing that poorer will get higher IRs? and richer pay less?
boxplot(ir ~ cut(loansData$Debt.To.Income.Ratio, breaks=seq(0,max(loansData$Debt.To.Income.Ratio, na.rm=T),max(loansData$Debt.To.Income.Ratio, na.rm=T)/20)))
# different cut
boxplot(ir ~ cut(loansData$Debt.To.Income.Ratio, breaks=10))

# there seems to be no relation with CB
boxplot(ir ~ cut(loansData$Revolving.CREDIT.Balance, breaks=20))

# this might be just reflection of the fact that 'poorer' people 
# have worse credit ratings, need more money and ask moore often for
# credit ratings, but look at how many outliers are for 0'th group
boxplot(ir ~ loansData$Inquiries.in.the.Last.6.Months)

# finally, IR may depend on the ammount requested
plot(y=loansData$Interest.Rate, x=loansData$Amount.Requested, log="xy")

# also Dept to income ratio may represent a linear relationship
plot(y=loansData$Interest.Rate, x=loansData$Debt.To.Income.Ratio, log="y")

# but there is no relationship with monthly income
plot(y=loansData$Interest.Rate, x=loansData$Monthly.Income, log="xy")
lm1 <- lm(log(ir) ~ log(loansData$Monthly.Income))
summary(lm1)

# but there might be with the Revolving credit balance
plot(y=log(loansData$Interest.Rate+1), x=log(loansData$Revolving.CREDIT.Balance+1))

lm1 <- lm(log(loansData$Interest.Rate+1) ~ log(loansData$Revolving.CREDIT.Balance+1))
summary(lm1)
```


## Modeling


## Fit a basic model relating IR to other vars, namely

 * FICO (maybe cut by quantiles)
 * Debt.To.Income.Ratio
 * Open.CREDIT.Lines [found not interesting]
 * Credit.Balance
 * Loan.Length
 * Ownership (but only for the category NONE?)
 * Purpose (especially: educational, car, renevable_energy)

TODO: look at relation between Credit.Balance ~ FICO

```{r, dependson="loansData"}

# FICO Good!
ir <- loansData$Interest.Rate
lmFico <- lm(ir ~ loansData$FICO.Range)
summary(lmFico)

# much bettter! select the median value as a base point
lmFico <- lm(ir ~ relevel(loansData$FICO.Range, ref="700-704"))
summary(lmFico)

# Debt.To.Income.Ratio Good!
lmDebRatio <- lm(ir ~ cut(loansData$Debt.To.Income.Ratio, breaks=20))
summary(lmDebRatio)


# maybe useful - but should be alredy inside FICO
lmOcr <- lm(ir ~ cut(loansData$Open.CREDIT.Lines, breaks=10))
summary(lmOcr)
confint(lmOcr)

# definitely incorporate Loan.Length
lmLen <- lm(ir ~ loansData$Loan.Length)
summary(lmLen)

# not sure, shall we?
levels(loansData$Home.Ownership)
lmHome <- lm(ir ~ loansData$Home.Ownership)
summary(lmHome)

# Revolving Credit balence: good!
lmBal <- lm(ir ~ log(loansData$Revolving.CREDIT.Balance + 1))
summary(lmBal)

# some categories of the Purpose are significant
# car, credit_card, debt_consolidation, house, moving
# other, small_business
levels(loansData$Loan.Purpose)
lmPurp <- lm(ir ~ loansData$Loan.Purpose)
summary(lmPurp)

# anova
anova(lmPurp)
# trying to use Tukey to see which combination is strong
#lmt <- aov(lmPurp)
#TukeyHSD(lmt)
```

## Plot the observed ('black') and fitted ('red') points

```{r, dependson="loansData"}

plot(loansData$FICO.Range, ir)
points(loansData$FICO.Range,lmFico$fitted.values,pch=19,col="red")


loans <- loansData
loans$FICO.Ordered <- relevel(loans$FICO.Range, ref="700-704")
loans$logBalance <- log10(loans$Revolving.CREDIT.Balance + 1)
loans$logIncome <- log10(loansData$Monthly.Income + 1)

ir <- loans$Interest.Rate

pdf(file="./figures/finalfigure.pdf", height=8, width=3*4)


par(mfrow=c(2,2))

lm1 <- lm(ir ~ loans$FICO)
summary(lm1)
confint(lm1)
plot(loans$FICO, ir, xlab="Credit Score",ylab="Interest Rate")
#points(lm1$fitted.values, col="red", pch=19)
abline(lm1, col="red", lw=3)

lm1 <- lm(ir ~ loans$Loan.Length)
summary(lm1)
confint(lm1)
plot(loans$Loan.Length, ir, xlab="Loan Length",ylab="Interest Rate")
abline(lm1, col="red", lw=3)
#points(lm1$fitted.values, col="red", pch=19)


lm1 <- lm(ir ~ loans$Amount.Requested)
summary(lm1)
confint(lm1)
plot(loans$Amount.Requested, ir, xlab="Amount requested",ylab="Interest Rate")
#points(lm1$fitted.values, col="red", pch=19)
abline(lm1, col="red", lw=3)

lm1 <- lm(ir ~ log10(loans$Revolving.CREDIT.Balance + 1))
summary(lm1)
confint(lm1)
plot(log10(loans$Revolving.CREDIT.Balance), ir, xlab="Revolving credit balance (log scale)",ylab="Interest Rate")
abline(lm1, col="red", lw=3)

lm1 <- lm(ir ~ loans$Open.CREDIT.Lines + loans$FICO.Ordered)
summary(lm1)
confint(lm1)
plot(loans$Open.CREDIT.Lines, ir, xlab="Credit inquiries (past 6 months)",ylab="Interest Rate")
#points(lm1$fitted.values, col="red", pch=19)
abline(lm1, col="red", lw=3)
dev.off()

lm1 <- lm(ir ~ log10(loans$Amount.Requested + 1))
summary(lm1)
confint(lm1)
plot(log10(loans$Revolving.CREDIT.Balance + 1), ir, xlab="Amount requested",ylab="Interest Rate")
abline(lm1, col="red", lw=3)



lm1 <- lm(ir ~ log10(loans$Revolving.CREDIT.Balance + 1))
summary(lm1)
confint(lm1)
plot(log10(loans$Revolving.CREDIT.Balance), ir)
abline(lm1, col="red", lw=3)
points(lm1$fitted.values, col="red", pch=19)

lm1 <- lm(ir ~ loans$FICO.Ordered + loans$logIncome)
summary(lm1)
confint(lm1)
plot(loans$logIncome, ir)
points(lm1$fitted.values, col="red", pch=19)

lm1 <- lm(ir ~ log10(loans$Revolving.CREDIT.Balance + 1))
summary(lm1)
confint(lm1)
plot(log10(loans$Revolving.CREDIT.Balance), ir, xlab="Revolving credit balance (log scale)",ylab="Interest Rate")
abline(lm1, col="red", lw=3)


plot(ir ~ loans$logIncome)
abline(lm1)

lm1 <- lm(ir ~ loans$logIncome + loans$Loan.Length + loans$Debt.To.Income.Ratio + loans$Revolving.CREDIT.Balance + loans$Open.CREDIT.Lines)
summary(lm1)
confint(lm1)

lm1 <- lm(ir ~ loans$FICO.Ordered + loans$Loan.Length + loans$Debt.To.Income.Ratio + loans$Revolving.CREDIT.Balance + loans$Open.CREDIT.Lines + loans$logIncome)
summary(lm1)
plot(ir)
```


## Try to fit the multi-variate model

```{r}
lmFinal <- lm(ir ~ loans$FICO + loans$Loan.Length + loans$logIncome + loans$Amount.Requested + log(loans$Revolving.CREDIT.Balance + 1))
summary(lmFinal)
confint(lmFinal)
#plot(lmFinal)
#plot(ir, lmFinal$fitted.values,pch=19,col=loans$Loan.Purpose)
#abline(c(lmFinal$coeff[1],lmFinal$coeff[2]),col="red",lwd=3)
#lines(ir,lmFinal$fitted,col="black",lwd=3)

#plot(lmFinal$fitted.values, ir)

# this looks like the jleeks plots, but i don't know how to 
# interpret them
plot(loans$Interest.Rate,lmFinal$residuals,pch=19, col=loans$FICO)
plot(loans$Interest.Rate,lmFinal$residuals,pch=19, col=loans$Loan.Length)
plot(loans$Interest.Rate,lmFinal$residuals,pch=19, col=loans$Amount.Requested)

```