Exploratory analysis
========================================================


Download the data, read the data in and save the raw data as an RDA file

```{r loansData, cache=TRUE}
getwd()
download.file("https://spark-public.s3.amazonaws.com/dataanalysis/loansData.csv",destfile="../../data/raw/loansData.csv",method="curl")
dateDownloaded <- date()
dateDownloaded
loansData <- read.csv('../../data/raw/loansData.csv')
save(loansData,dateDownloaded,file="../../data/raw/loansData.rda")
```

### Fix incorrect types
```{r dependson="loansData"}
loansData$Interest.Rate <- sapply(loansData$Interest.Rate, function(x) as.numeric(sub("%", "", x)))
loansData$Debt.To.Income.Ratio <- sapply(loansData$Debt.To.Income.Ratio, function(x) as.numeric(sub("%", "", x)))
```

### Look at the data set

```{r}
str(loansData)
sapply(loansData, summary)
```

### Find out about missing values and remove the incomplete
```{r, dependson="loansData"}
sum(is.na(loansData))
loansData <- loansData[complete.cases(loansData),]
```


## Exploratory analysis

### Make some univariate plots/summaries

```{r, dependson="loansData"}
loansData$FICO <- sapply(as.character(loansData$FICO.Range), function(s) as.numeric(substr(s, 0, 3)))

# interest rate seems interesting
hist(loansData$Interest.Rate, breaks=100)
# as if there were two distributions...
hist(log10(loansData$Interest.Rate), breaks = 100, prob=T)

# this is interesting, there is a 'required' minimum
hist(loansData$FICO, breaks=100)

# is this increasing lineary? it seems like...
quantile(loansData$FICO, na.rm=T, probs=seq(0, 1, 0.05))

# investors obviously prefer borrowers who don't have too
# high ratio of indebtness, but it is not too strong
hist(loansData$Debt.To.Income.Ratio)

# and few credit lines
hist(loansData$Open.CREDIT.Lines, breaks=100)
hist(as.numeric(cut(loansData$Open.CREDIT.Lines, breaks=seq(0,30,1))))


# my understanding is that only borrowers who do not
# already own too much, are welcomed
# TODO: look at the relationship between CB ~ FICO
quantile(loansData$Revolving.CREDIT.Balance, na.rm=T, probs=seq(0, 1, 0.05))
hist(loansData$Revolving.CREDIT.Balance, breaks=100)
# needs transformation...
hist(log(loansData$Revolving.CREDIT.Balance), breaks=100)

```

So it seems that FICO must be > 650 for the loan to be approved. And we have got some terms to check for relations

### Fix the distribution scales
```{r, dependson="loansData"}

loansData$logIncome <- log10(loansData$Monthly.Income + 1)
loansData$logRevCredit <- log(loansData$Revolving.CREDIT.Balance + 1)
# I want the median to be intercepted first
loansData$FICO.Ordered <- relevel(loansData$FICO.Range, ref="700-704")
```

### Check out relationship with Interest Rate (IR)
```{r, dependson="loansData"}
ir <- loansData$Interest.Rate
par(las=3)
par(mar=c(6,2,1,1))
boxplot(ir ~ loansData$Loan.Purpose )
# it seems significant, but in the end i decided not to use it
anova(lm(ir ~ loansData$Loan.Purpose))


boxplot(ir ~ loansData$Home.Ownership)
anova(lm(ir ~ loansData$Home.Ownership))

# a linear relationn between IR an FICO?
boxplot(ir ~ loansData$FICO.Range)

# no relation between IR and Employment length
boxplot(ir ~ loansData$Employment.Length)

# IR clearly depends on the length (which is logical, but
# it is interesting that shorter terms have lower IR)
boxplot(ir ~ loansData$Loan.Length)

# am i seeing that poorer will get poorer? and richer pay less?
boxplot(ir ~ cut(loansData$Debt.To.Income.Ratio, breaks=seq(0,max(loansData$Debt.To.Income.Ratio, na.rm=T),max(loansData$Debt.To.Income.Ratio, na.rm=T)/20)))

# different cut
boxplot(ir ~ cut(loansData$Debt.To.Income.Ratio, breaks=10))


# finally, IR may depend on the ammount requested
plot(y=loansData$Interest.Rate, x=loansData$Amount.Requested, log="xy")

# also Dept to income ratio may represent a linear relationship
plot(y=loansData$Interest.Rate, x=loansData$Debt.To.Income.Ratio, log="y")

# but there is no relationship with monthly income
plot(y=loansData$Interest.Rate, x=loansData$Monthly.Income, log="xy")
lm1 <- lm(log(ir) ~ log(loansData$Monthly.Income))
summary(lm1)

# but there might be with the Revolving credit balance
plot(y=log(loansData$Interest.Rate+1), x=log(loansData$Revolving.CREDIT.Balance+1))
lm1 <- lm(log(loansData$Interest.Rate+1) ~ log(loansData$Revolving.CREDIT.Balance+1))
summary(lm1)
```


## Modeling


## Fit a basic model relating IR to other vars, namely

 * FICO (maybe cut by quantiles)
 * Debt.To.Income.Ratio
 * Open.CREDIT.Lines [found not interesting]
 * Credit.Balance
 * Loan.Length
 * Ownership (but only for the category NONE?)
 * Purpose (especially: educational, car, renevable_energy)



```{r, dependson="loansData"}

# FICO is the base
ir <- loansData$Interest.Rate
lmFico <- lm(ir ~ loansData$FICO.Ordered)
summary(lmFico)


# Debt.To.Income.Ratio might be useful
lmDebRatio <- lm(ir ~ loansData$Debt.To.Income.Ratio)
summary(lmDebRatio)
plot(loansData$Debt.To.Income.Ratio, ir)
points(lmDebRatio$fitted.values, col="red")
abline(lmDebRatio, col="red", lw=3)


# definitely incorporate Loan.Length
lmLen <- lm(ir ~ loansData$Loan.Length)
summary(lmLen)

# not sure, probably going to ignore ownership
levels(loansData$Home.Ownership)
lmHome <- lm(ir ~ loansData$Home.Ownership)
summary(lmHome)

# Revolving Credit balence: seems to matter (but it has weird shape)
lmBal <- lm(ir ~ loansData$logRevCredit)
plot(loansData$logRevCredit, ir)
points(lmBal$fitted.values, col="red")
abline(lmBal, col="red", lw=3)
summary(lmBal)

# some categories of the Purpose are significant
# car, credit_card, debt_consolidation, house, moving
# other, small_business
# but I will ignore it....
levels(loansData$Loan.Purpose)
lmPurp <- lm(ir ~ loansData$Loan.Purpose)
summary(lmPurp)


```

## Try to fit the multi-variate model

```{r}
ir <- loansData$Interest.Rate
# first, i thought about using Debt.to.Income.Ratio instead of
# logRevCredit, but it seems that logRevCredit has more impact
lmFinal <- lm(ir ~ loansData$FICO + loansData$Loan.Length + loansData$logIncome + loansData$Amount.Requested + loansData$logRevCredit)
summary(lmFinal)
confint(lmFinal)


# this looks like plots from example, but i don't know how to 
# interpret them...
plot(loansData$Interest.Rate,lmFinal$residuals,pch=19, col=loansData$FICO)
plot(loansData$Interest.Rate,lmFinal$residuals,pch=19, col=loansData$Loan.Length)
plot(loansData$Interest.Rate,lmFinal$residuals,pch=19, col=loansData$Amount.Requested)
plot(loansData$Interest.Rate,lmFinal$residuals,pch=19, col=loansData$Debt.To.Income.Ratio)

```

## Plot the influence of different terms

```{r, dependson="loansData"}


pdf(file="../../figures/finalfigure.pdf", height=8, width=3*4)


par(mfrow=c(2,2))

lm1 <- lm(ir ~ loansData$FICO)
summary(lm1)
confint(lm1)
plot(loansData$FICO, ir, xlab="Credit Score",ylab="Interest Rate")
#points(lm1$fitted.values, col="red", pch=19)
abline(lm1, col="red", lw=3)

lm1 <- lm(ir ~ loansData$Loan.Length)
summary(lm1)
confint(lm1)
plot(loansData$Loan.Length, ir, xlab="Loan Length",ylab="Interest Rate")
abline(lm1, col="red", lw=3)
#points(lm1$fitted.values, col="red", pch=19)


lm1 <- lm(ir ~ loansData$Amount.Requested)
summary(lm1)
confint(lm1)
plot(loansData$Amount.Requested, ir, xlab="Amount requested",ylab="Interest Rate")
#points(lm1$fitted.values, col="red", pch=19)
abline(lm1, col="red", lw=3)

lm1 <- lm(ir ~ loansData$logRevCredit)
summary(lm1)
confint(lm1)
plot(loansData$logRevCredit, ir, xlab="Revolving credit balance (log scale)", ylab="Interest Rate")
abline(lm1, col="red", lw=3)

```


